Bonomi et al. introduce a simple, general reweighting scheme for WTMetaD that recovers unbiased Boltzmann statistics of any observable—starting from the same key identity defining $P(\mathbf r,t)$ (Eq. \ref{eq:metad_sampled_prob}), the definition of $c(t)$ (Eq. \ref{eq:coft}) and the long-time limit of the bias constructed with WTmetaD: $V(\xi,t\to\infty)=-\Delta T/(\Delta T+T)\,F(\xi)$, Bonomi et al. develop a reweighting approach that circumvents the need of computing $c(t)$. By differentiating Eq. \ref{eq:metad_sampled_prob} for small time intervals $\Delta t$,  an evolution equation that eliminates $c(t)$ is derived: 
\begin{equation}
p(\mathbf r,t+\Delta t)=e^{-\beta\,[\dot V(\xi(\mathbf r),t)-\langle \dot V(\xi,t)\rangle]\Delta t}\,p(\mathbf r,t)
\label{eq:propagator}
\end{equation}
where $\dot c(t)=-\langle \dot V(\xi,t)\rangle$ and the average is over the biased distribution at time $t$. For WTMetaD with Gaussian depositions, these results lead to a practical reweighting algorithm based on three steps. (i) Accumulate a joint histogram $N_t(\xi,f)$ for a target variable $f(\mathbf r)$ between Gaussian updates; (ii) at each update, compute $\dot V$ and $\dot c$ using the current accumulated histogram, and evolve $N_t$ using Eq. \ref{eq:propagator}; (iii) reconstruct the unbiased distribution of $f(\mathbf{r})$ as
\begin{equation}
p(f)=\frac{\sum_{\xi} e^{+\beta V(\xi,t)}\,N_t(\xi,f)}{\sum_{\xi,f} e^{+\beta V(\xi,t)}\,N_t(\xi,f)}.
\end{equation}
The method is lightweight as it does not require any a posteriori calculation of the total energy, it works in post-processing or (in principle) on-the-fly, and it converges efficiently as shown in  Refs. \cite{gimondi2018building,Marinova2019}.
